---
title: "Old Chicago Main Post Office: Difference-in-Differences Analysis"
author: ""
date: "April 13, 2026"
output:
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: true
    fig_width: 7
    fig_height: 4
    fig_caption: true
    latex_engine: xelatex
geometry: margin=1in
fontsize: 11pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center"
)

required_packages <- c("data.table", "ggplot2", "fixest", "stringr", "here", "scales")
new_packages <- required_packages[!(required_packages %in% installed.packages()[, "Package"])]
if (length(new_packages)) install.packages(new_packages)

library(data.table)
library(ggplot2)
library(fixest)
library(stringr)
library(here)
library(scales)
```

```{r file-paths}
treat_file       <- here("treatment_0_500.csv")
control_file     <- here("control_800_1300.csv")
post_office_file <- here("Assessor_-_Assessed_Values_20260401 (1).csv")
south1_file      <- here("South Chicago_1_Assessed_Values.csv")
south2_file      <- here("South Chicago_2_Assessed_Values.csv")
west1_file       <- here("West Chicago_1_Assessed_Values.csv")
west2_file       <- here("West Chicago_2_Assessed_Values.csv")
```

```{r load-parcels}
treat   <- fread(treat_file)
control <- fread(control_file)

treat   <- treat[,   .(keypin, bldgsf, usedesc)]
control <- control[, .(keypin, bldgsf, usedesc)]

treat[,   group := "treatment"]
control[, group := "control"]

sample_parcels <- rbindlist(list(treat, control), fill = TRUE)

sample_parcels[, pin    := str_pad(as.character(keypin), width = 14, pad = "0")]
sample_parcels[, bldgsf := as.numeric(bldgsf)]

sample_parcels <- sample_parcels[!is.na(bldgsf) & bldgsf > 0]
sample_parcels <- unique(sample_parcels[, .(pin, bldgsf, usedesc, group)], by = "pin")

dup_pins <- sample_parcels[, .N, by = pin][N > 1]
if (nrow(dup_pins) > 0) stop("Some PINs appear in both treatment and control groups.")
```

```{r load-assessor}
assessor <- rbindlist(list(
  fread(south1_file), fread(south2_file),
  fread(west1_file),  fread(west2_file)
), fill = TRUE)

assessor <- assessor[, .(pin, tax_year, township_name, class, certified_tot)]
assessor[, pin          := str_pad(as.character(pin), width = 14, pad = "0")]
assessor[, certified_tot := as.numeric(gsub("[$,]", "", certified_tot))]
assessor <- assessor[!is.na(certified_tot)]

assessor_small <- assessor[pin %in% sample_parcels$pin]

panel <- merge(assessor_small, sample_parcels, by = "pin", all.x = TRUE)
panel <- panel[tax_year >= 1999 & tax_year <= 2025]
panel <- panel[certified_tot > 0]
```

```{r cpi}
# CPI-U annual averages, U.S. City Average
# Source: Bureau of Labor Statistics. Base year: 2025.
cpi <- data.table(
  tax_year = 1999:2025,
  cpi = c(
    168.400, 173.800, 178.300, 181.200, 184.500,
    188.600, 194.300, 198.300, 204.818, 212.536,
    209.995, 212.870, 218.684, 222.005, 224.545,
    228.468, 227.792, 229.302, 233.611, 237.706,
    241.181, 243.873, 254.159, 273.356, 282.450,
    292.225, 301.232
  )
)

base_cpi <- cpi[tax_year == 2025, cpi][1]

panel <- merge(panel, cpi, by = "tax_year", all.x = TRUE)
panel[, real_certified_tot := certified_tot * (base_cpi / cpi)]
```

```{r value-psf}
panel[, value_psf := real_certified_tot / bldgsf]
panel <- panel[!is.na(value_psf) & is.finite(value_psf) & value_psf >= 1]

p99          <- panel[, quantile(value_psf, 0.99, na.rm = TRUE)]
panel_main   <- panel[value_psf <= p99]
```

```{r did-setup}
panel_main[, treatment := ifelse(group == "treatment", 1, 0)]
panel_main[, post      := ifelse(tax_year >= 2019, 1, 0)]

# Main specification: all years
did_data_full <- panel_main[value_psf > 0]
did_data_full[, log_value_psf := log(value_psf)]
did_data_full[, year_rel      := tax_year - 2019]

# Robustness specification: exclude 2016-2018 transition years
did_data_clean <- panel_main[!tax_year %in% c(2016, 2017, 2018) & value_psf > 0]
did_data_clean[, log_value_psf := log(value_psf)]
did_data_clean[, year_rel      := tax_year - 2019]
```

```{r theme}
# Shared plot theme
theme_po <- function() {
  theme_minimal(base_size = 11) +
    theme(
      plot.title    = element_text(face = "bold", size = 11),
      plot.subtitle = element_text(size = 9, color = "gray40"),
      legend.position = "bottom"
    )
}

# Shared annotation layers
vlines <- list(
  geom_vline(xintercept = 2010, linetype = "dashed", color = "gray50"),
  geom_vline(xintercept = 2019, linetype = "dashed", color = "black")
)

transition_band <- annotate(
  "rect", xmin = 2016, xmax = 2018,
  ymin = -Inf, ymax = Inf,
  alpha = 0.15, fill = "gray70"
)
```

\newpage

# Post Office Building Value Trajectory

```{r post-office-series}
post_office_pin    <- "17161300010000"
post_office_series <- fread(post_office_file)

post_office_series <- post_office_series[, .(pin, tax_year, certified_tot)]
post_office_series[, pin          := str_pad(as.character(pin), width = 14, pad = "0")]
post_office_series[, certified_tot := as.numeric(gsub("[$,]", "", certified_tot))]
post_office_series <- post_office_series[
  pin == post_office_pin & !is.na(certified_tot)
][order(tax_year)]

post_office_series <- merge(post_office_series, cpi, by = "tax_year", all.x = TRUE)
post_office_series[, real_certified_tot := certified_tot * (base_cpi / cpi)]
```

```{r fig-post-office, fig.cap="Real assessed value of the Old Chicago Main Post Office parcel, 1999-2025. Shaded band marks 2016-2018 transition period."}
ggplot(post_office_series, aes(x = tax_year, y = real_certified_tot / 1e6)) +
  transition_band +
  vlines +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  scale_y_continuous(labels = label_number(accuracy = 1)) +
  labs(
    title = "Old Chicago Main Post Office: Real Assessed Value (1999-2025)",
    subtitle = paste(
      "Shaded band = 2016-2018 transition years.",
      "Dashed lines = 2010 baseline and 2019 treatment onset."
    ),
    x = "Tax Year",
    y = "Real assessed value (USD millions, 2025 dollars)"
  ) +
  theme_po()
```

\newpage

# Parallel Trends

## Formal pre-period: 1999-2015

```{r pre-trends-strict}
pre_trends_strict <- panel_main[tax_year <= 2015][
  ,
  .(
    avg_value_psf     = mean(value_psf, na.rm = TRUE),
    log_avg_value_psf = log(mean(value_psf, na.rm = TRUE)),
    n = .N
  ),
  by = .(tax_year, group)
][order(group, tax_year)]
```

```{r fig-pre-levels, fig.cap="Average real assessed value per sq ft, treatment vs control, 1999-2015. Dashed line marks 2010 baseline."}
ggplot(pre_trends_strict, aes(x = tax_year, y = avg_value_psf, color = group)) +
  geom_vline(xintercept = 2010, linetype = "dashed", color = "gray50") +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  labs(
    title    = "Pre-Treatment Trends: Real Assessed Value per Sq Ft (1999-2015)",
    subtitle = "Dashed line marks 2010 baseline",
    x = "Tax Year",
    y = "Avg real assessed value per sq ft",
    color = "Group"
  ) +
  theme_po()
```

```{r fig-pre-logs, fig.cap="Log average real assessed value per sq ft, treatment vs control, 1999-2015."}
ggplot(pre_trends_strict, aes(x = tax_year, y = log_avg_value_psf, color = group)) +
  geom_vline(xintercept = 2010, linetype = "dashed", color = "gray50") +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  labs(
    title    = "Pre-Treatment Trends: Log Real Assessed Value per Sq Ft (1999-2015)",
    subtitle = "Dashed line marks 2010 baseline",
    x = "Tax Year",
    y = "Log avg real assessed value per sq ft",
    color = "Group"
  ) +
  theme_po()
```

## Extended pre-period including transition years: 1999-2018

```{r pre-trends-hist}
pre_trends_hist <- panel_main[tax_year <= 2018][
  ,
  .(
    avg_value_psf     = mean(value_psf, na.rm = TRUE),
    log_avg_value_psf = log(mean(value_psf, na.rm = TRUE)),
    n = .N
  ),
  by = .(tax_year, group)
][order(group, tax_year)]
```

```{r fig-hist-levels, fig.cap="Average real assessed value per sq ft including 2016-2018 transition years. Shaded band marks transition period."}
ggplot(pre_trends_hist, aes(x = tax_year, y = avg_value_psf, color = group)) +
  transition_band +
  geom_vline(xintercept = 2010, linetype = "dashed", color = "gray50") +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  labs(
    title    = "Pre-2019 Trends: Real Assessed Value per Sq Ft (1999-2018)",
    subtitle = paste(
      "Shaded band = 2016-2018 transition years.",
      "Dashed line = 2010 baseline."
    ),
    x = "Tax Year",
    y = "Avg real assessed value per sq ft",
    color = "Group"
  ) +
  theme_po()
```

```{r fig-hist-logs, fig.cap="Log average real assessed value per sq ft including 2016-2018 transition years."}
ggplot(pre_trends_hist, aes(x = tax_year, y = log_avg_value_psf, color = group)) +
  transition_band +
  geom_vline(xintercept = 2010, linetype = "dashed", color = "gray50") +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  labs(
    title    = "Pre-2019 Trends: Log Real Assessed Value per Sq Ft (1999-2018)",
    subtitle = paste(
      "Shaded band = 2016-2018 transition years.",
      "Dashed line = 2010 baseline."
    ),
    x = "Tax Year",
    y = "Log avg real assessed value per sq ft",
    color = "Group"
  ) +
  theme_po()
```

\newpage

# Full-Period Descriptive Trends (1999-2025)

```{r full-trends}
yearly_means_full <- did_data_full[
  ,
  .(
    avg_value_psf     = mean(value_psf,     na.rm = TRUE),
    avg_log_value_psf = mean(log_value_psf, na.rm = TRUE),
    n = .N
  ),
  by = .(tax_year, group, treatment)
][order(group, tax_year)]
```

```{r fig-full-levels, fig.cap="Full-period average real assessed value per sq ft. Shaded band = 2016-2018 transition years. Dashed lines = 2010 baseline and 2019 treatment onset."}
ggplot(yearly_means_full, aes(x = tax_year, y = avg_value_psf, color = group)) +
  transition_band +
  vlines +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  labs(
    title    = "Treatment vs Control: Real Assessed Value per Sq Ft (1999-2025)",
    subtitle = paste(
      "Shaded band = 2016-2018 transition years.",
      "Dashed lines = 2010 baseline and 2019 treatment onset."
    ),
    x = "Tax Year",
    y = "Avg real assessed value per sq ft",
    color = "Group"
  ) +
  theme_po()
```

```{r fig-full-logs, fig.cap="Full-period log average real assessed value per sq ft."}
ggplot(yearly_means_full, aes(x = tax_year, y = avg_log_value_psf, color = group)) +
  transition_band +
  vlines +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  labs(
    title    = "Treatment vs Control: Log Real Assessed Value per Sq Ft (1999-2025)",
    subtitle = paste(
      "Shaded band = 2016-2018 transition years.",
      "Dashed lines = 2010 baseline and 2019 treatment onset."
    ),
    x = "Tax Year",
    y = "Log avg real assessed value per sq ft",
    color = "Group"
  ) +
  theme_po()
```

\newpage

# Difference-in-Differences Results

## Two-Way Fixed Effects Models

```{r did-twfe}
did_fe_log_full <- feols(
  log_value_psf ~ treatment:post | pin + tax_year,
  cluster = ~pin,
  data    = did_data_full
)

did_fe_log_clean <- feols(
  log_value_psf ~ treatment:post | pin + tax_year,
  cluster = ~pin,
  data    = did_data_clean
)

cat("Main specification (all years):\n")
summary(did_fe_log_full)

cat("\nRobustness check (excluding 2016-2018):\n")
summary(did_fe_log_clean)
```

## Coefficient Comparison

```{r model-compare}
model_compare <- data.table(
  Specification = c("Main (all years)", "Robustness (excl. 2016-2018)"),
  Coef  = c(coef(did_fe_log_full)["treatment:post"],
             coef(did_fe_log_clean)["treatment:post"]),
  SE    = c(fixest::se(did_fe_log_full)["treatment:post"],
             fixest::se(did_fe_log_clean)["treatment:post"]),
  Pval  = c(fixest::pvalue(did_fe_log_full)["treatment:post"],
             fixest::pvalue(did_fe_log_clean)["treatment:post"]),
  N     = c(nobs(did_fe_log_full), nobs(did_fe_log_clean))
)

model_compare[, CI_low  := Coef - 1.96 * SE]
model_compare[, CI_high := Coef + 1.96 * SE]

cols <- c("Coef", "SE", "CI_low", "CI_high")
model_compare[, (cols) := lapply(.SD, round, 4), .SDcols = cols]
model_compare[, Pval := round(Pval, 4)]

print(model_compare)
```

\newpage

# Event Study

The event study uses the robustness sample (2016-2018 excluded) with year $-4$
(2015) as the reference period. Pre-period coefficients near zero support the
parallel trends assumption.

```{r event-study, fig.cap="Event study coefficients for log real assessed value per sq ft. Reference year is 2015 (year -4). Bands show 95% confidence intervals."}
es_model <- feols(
  log_value_psf ~ i(year_rel, treatment, ref = -4) | pin + tax_year,
  cluster = ~pin,
  data    = did_data_clean
)

summary(es_model)

iplot(
  es_model,
  main      = "Event Study: Log Real Assessed Value per Sq Ft",
  xlab      = "Years relative to treatment onset (2019)",
  ylab      = "Coefficient (log points)",
  col       = "steelblue",
  zero.line = TRUE
)
```

\newpage

# Appendix: OLS Models

Included for reference only. The TWFE results in Section 4 are the preferred
specification; OLS conflates within- and between-unit variation.

```{r appendix-ols}
cat("Main specification:\n")
did_model_level_full <- lm(value_psf    ~ treatment * post, data = did_data_full)
did_model_log_full   <- lm(log_value_psf ~ treatment * post, data = did_data_full)
summary(did_model_level_full)
summary(did_model_log_full)

cat("\nRobustness (excluding 2016-2018):\n")
did_model_level_clean <- lm(value_psf    ~ treatment * post, data = did_data_clean)
did_model_log_clean   <- lm(log_value_psf ~ treatment * post, data = did_data_clean)
summary(did_model_level_clean)
summary(did_model_log_clean)
```
